Effectiveness of visual test in linear regression diagnositics

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.Rmd
reorder code

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. Information about gender, age, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

Clearly, omitted-variable bias will present since the null model leaves out the higher order term. Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(a = -1, b = 1)$, (2) $N(\mu = 0, \sigma = 0.3)$, (3) $lognormal(\mu = 0, \sigma = 0.6)/3$ and (4) $u\{a = 1, b = 5\}$. A summary of the parameters used in this experiment is given in [table ref here]

The values of $j$ was chosen so that visually different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

@@

The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated and the estimated power curve would be smooth and continuous. Four different distribution were used to genearte $\boldsymbol{X}_raw$.

@@

The uniform and the normal distribution are symmetric and commonly assumed in statistical model. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

VI_MODEL$plot(ori_dat) + ggtitle(quote(U(-1,1))) -> p1
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(N(0,0.3))) -> p2
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(lognormal(0,0.6)/3)) -> p3
VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(U~"{1,5}")) -> p4

@@

Three replications are made for each of the parameter values shown in [table ref here] and thus, 192 different lineups. For each lineup, the null model was fit to the actual dataset to obtain residuals and parameter estimates. The actual data plot was drawn as standard residual plot with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The actual data plot was randomly placed among these null dadta plots to produce the lineup. [figure ref here] is an example of one of these lineups. It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$. For this lineup, x out of k observers picked the actual data plot. Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value.

To satisfy the plan described above, we needed to recruit at least 160 subjects.

@@

## Data cleaning

@@

# Demographic summary

>>>>

As suggested by many previous studies [many studies refs here], it is reasonable for an subject to evaluate a block of 20 lineups. Subjects for both two experiments were recruited from Prolific web service [ref here]. They were asked to select one or more plots that are mostly different from others, provide a reason for their selections, and express how different they think the selected plots are from others. If there was no noticeable difference between plots in a lineup, subject could choose to select zero lineups without providing the reason. Information about gender, age, education, and previous experience in visual experiment were also collected. No subject was shown the same lineup twice.

@@

Clearly, omitted-variable bias will present since the null model leaves out the higher order term. Data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $\boldsymbol{X}_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

The values of $j$ was chosen so that different shapes of non-linearity are included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

@@

The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated and the estimated power curve would be smooth and continuous.

@@

Four different distribution were used to generate $\boldsymbol{X}_{raw}$. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

VI_MODEL$plot(ori_dat) + ggtitle(quote(U(-1,1))) -> p1
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(N(0,0.3^2))) -> p2
VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(lognormal(0,0.6^2)/3)) -> p3
VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle(quote(U~"{1,5}")) -> p4

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in 192 different lineups. For each lineup, the actual data plot was drawn as standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. Figure \ref{fig:example-lineup} is an example of one of these lineups. It was produced by using $n = 300$, $j = 6$, $\sigma = 0.5$ and $N(0.0.3^2)$ as the distribution of $X_{raw}$. The actual data plot location is four. All five subjects correctly identified the actual data plot for this lineup.

```{r results='asis'}
data.frame(n = c(50, 100, 300, ""),
           j = c(2, 3, 6, 18),
           sigma = c(0.5, 1, 2, 4),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$\\\\sigma$", .) %>%
  gsub("sym3", "$n$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$. For this lineup, x out of k observers picked the actual data plot. Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value. -->

In addition, each lineup is evaluated by five different subjects to provide reasonable estimates of the visual p-value. Thus, $192 \times 3 \times 5 / (20 - 2) = 160$ subjects were recruited to satisfy the experimental design described above.

@@

## Data preprocessing

Subjects recruited from Prolific received a fixed payment for submitting quality responses. However, some subjects will try to maximize their earnings for minimum effort. If a subject objectively demonstrated clear low-effort throughout the experiment, i.e., failed both attention checks, we rejected the submission and the subject would not be paid.

@@

## Demographic summary

### 2022-07-28 paper_comparison/paper_comparison/paper_comparison.tex
reorder code

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%

@@

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. They were asked to select one or more plots that are mostly
different from others, provide a reason for their selections, and
express how different they think the selected plots are from others.
Information about gender, age, education, and previous experience in
visual experiment were also collected. No subject was shown the same
lineup twice.

@@

Clearly, omitted-variable bias will present since the null model leaves
out the higher order term. Data were simulated using four different
order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three
different sample sizes (n = 50, 100, 300) , four different standard
deviations of the error (\(\sigma\) = 0.5, 1, 2, 4) and four different
distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(a = -1, b = 1)\), (2)
\(N(\mu = 0, \sigma = 0.3)\), (3) \(lognormal(\mu = 0, \sigma = 0.6)/3\)
and (4) \(u\{a = 1, b = 5\}\). A summary of the parameters used in this
experiment is given in {[}table ref here{]}

The values of \(j\) was chosen so that visually different shapes of
non-linearity are included in the residual plot. These include ``U''
shape, ``S'' shape, ``M'' shape and ``Triple-U'' shape.

@@

The range of \(\sigma\), which is a factor controlling the strength of
the signal, was chosen so that different difficulty levels of lineups
were generated and the estimated power curve would be smooth and
continuous. Four different distribution were used to genearte
\(\boldsymbol{X}_raw\).

@@

The uniform and the normal distribution are symmetric and commonly
assumed in statistical model. The adjusted log-normal distribution
provides skewed density. And the discrete uniform distribution provides
discreteness in residual plot, which could enrich the pool of visual
patterns.

@@

Three replications are made for each of the parameter values shown in
{[}table ref here{]} and thus, 192 different lineups. For each lineup,
the null model was fit to the actual dataset to obtain residuals and
parameter estimates. The actual data plot was drawn as standard residual
plot with raw residuals on the y-axis and fitted values on the x-axis.
The 19 null datasets were generated by the residual rotation technique,
and plotted in the same way. The actual data plot was randomly placed
among these null dadta plots to produce the lineup. {[}figure ref
here{]} is an example of one of these lineups. It was generated for
\(n = 1\), \(\sigma = 1\), \(j = 1\), \(X_{raw} \sim N\). For this
lineup, x out of k observers picked the actual data plot. Each lineup is
evaluated by 5 different subjects to provide reasonable estimates of the
p-value.

To satisfy the plan described above, we needed to recruit at least 160
subjects.

@@

\hypertarget{data-cleaning}{%
\subsection{Data cleaning}\label{data-cleaning}}

@@

@@

\hypertarget{demographic-summary}{%
\section{Demographic summary}\label{demographic-summary}}

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}

>>>>

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

As suggested by many previous studies {[}many studies refs here{]}, it
is reasonable for an subject to evaluate a block of 20 lineups. Subjects
for both two experiments were recruited from Prolific web service {[}ref
here{]}. They were asked to select one or more plots that are mostly
different from others, provide a reason for their selections, and
express how different they think the selected plots are from others. If
there was no noticeable difference between plots in a lineup, subject
could choose to select zero lineups without providing the reason.
Information about gender, age, education, and previous experience in
visual experiment were also collected. No subject was shown the same
lineup twice.

@@

Clearly, omitted-variable bias will present since the null model leaves
out the higher order term. Data were simulated using four different
order of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three
different sample sizes (n = 50, 100, 300) , four different standard
deviations of the error (\(\sigma\) = 0.5, 1, 2, 4) and four different
distribution of \(\boldsymbol{X}_{raw}\): (1) \(U(-1, 1)\), (2)
\(N(0, 0.3^2)\), (3) \(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A
summary of the parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

The values of \(j\) was chosen so that different shapes of non-linearity
are included in the residual plot. These include ``U'' shape, ``S''
shape, ``M'' shape and ``Triple-U'' shape.

@@

The range of \(\sigma\), which is a factor controlling the strength of
the signal, was chosen so that different difficulty levels of lineups
were generated and the estimated power curve would be smooth and
continuous.

@@

Four different distribution were used to generate
\(\boldsymbol{X}_{raw}\). The uniform and the normal distribution are
symmetric and commonly assumed in statistical models. The adjusted
log-normal distribution provides skewed density. And the discrete
uniform distribution provides discreteness in residual plot, which could
enrich the pool of visual patterns.

@@

Three replications are made for each of the parameter values shown in
Table \ref{tab:parameter-table} resulting in 192 different lineups. For
each lineup, the actual data plot was drawn as standard residual plot of
the null model with raw residuals on the y-axis and fitted values on the
x-axis. The 19 null datasets were generated by the residual rotation
technique, and plotted in the same way. The lineup consisted of 20
residual plots with one randomly placed actual data plot. Figure
\ref{fig:example-lineup} is an example of one of these lineups. It was
produced by using \(n = 300\), \(j = 6\), \(\sigma = 0.5\) and
\(N(0.0.3^2)\) as the distribution of \(X_{raw}\). The actual data plot
location is four. All five subjects correctly identified the actual data
plot for this lineup.

\begin{table}

\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
\centering
\begin{tabular}[t]{lrrl}
\toprule
\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
50 & 2 & 0.5 & $U(-1, 1)$\\
100 & 3 & 1.0 & $N(0, 0.3^2)$\\
300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
 & 18 & 4.0 & $U\{1, 5\}$\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{paper_comparison_files/figure-latex/example-lineup-1.pdf}
\caption{Example lineup \label{fig:example-lineup}}
\end{figure}

In addition, each lineup is evaluated by five different subjects to
provide reasonable estimates of the visual p-value. Thus,
\(192 \times 3 \times 5 / (20 - 2) = 160\) subjects were recruited to
satisfy the experimental design described above.

@@

\hypertarget{data-preprocessing}{%
\subsection{Data preprocessing}\label{data-preprocessing}}

Subjects recruited from Prolific received a fixed payment for submitting
quality responses. However, some subjects will try to maximize their
earnings for minimum effort. If a subject objectively demonstrated clear
low-effort throughout the experiment, i.e., failed both attention
checks, we rejected the submission and the subject would not be paid.

@@

\hypertarget{demographic-summary}{%
\subsection{Demographic summary}\label{demographic-summary}}

@@
